---
title: "Econometrics_5300_Data_Anlalysis"
author: "Andrew Zinkan"
date: "8/1/2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes
<!-- fill na with zero -->
<!-- master_df[is.na(master_df)] <- 0 -->


## RESEARCH QUESTION:

The College Scorecard was released at the start of September 2015. Among colleges that predominantly grant bachelor’s degrees, did the release of the Scorecard shift student interest to high-earnings colleges relative to low-earnings ones (as proxied by Google searches for keywords associated with those colleges)?

REQ
- You will need to produce at least one regression and one graph for your analysis, and explain them.


```{}
# Load initial packages
library(tidyverse)
library(ggplot2)
library(dplyr)
library(vtable)
library(fixest)
```

### Load Raw Data
```{}
# Striping the Nulls to NA Values
DataDictRaw <-  read.csv("Data_Exploration/CollegeScorecardDataDictionary.csv", na.strings = c("","NA","NULL",NULL))
IDNameLinkRaw <- read.csv("Data_Exploration/id_name_link.csv", na.strings = c("","NA","NULL",NULL))
MostRecentRaw <- read.csv("Data_Exploration/Most_Recent_Cohorts_Scorecard_Elements.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToFinishRaw <- read.csv("Data_Exploration/trends_up_to_finish.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToInter_1_Raw <- read.csv("Data_Exploration/trends_up_to_inter_1.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToInter_2_Raw <- read.csv("Data_Exploration/trends_up_to_inter_2.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToInter_3_Raw <- read.csv("Data_Exploration/trends_up_to_inter_3.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToInter_4_Raw <- read.csv("Data_Exploration/trends_up_to_inter_4.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToInter_5_Raw <- read.csv("Data_Exploration/trends_up_to_inter_5.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToInter_6_Raw <- read.csv("Data_Exploration/trends_up_to_inter_6.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToUmRaw <- read.csv("Data_Exploration/trends_up_to_UM.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToUphoenixRaw <- read.csv("Data_Exploration/trends_up_to_UPhoenix.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToUtRaw <- read.csv("Data_Exploration/trends_up_to_UT.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToUtmbRaw <- read.csv("Data_Exploration/trends_up_to_UTMB.csv", na.strings = c("","NA","NULL",NULL))
TrendsUpToYorktowneRaw <- read.csv("Data_Exploration/trends_up_to_Yorktowne.csv", na.strings = c("","NA","NULL",NULL))
```

## Overview of Data

```{}
summary(DataDictRaw)
```
Notable Observations

- 1295 unique different variables
- Full Data Dict Available In Apedex()

``` {}
summary(IDNameLinkRaw)

```
Data set holds unique ids for the universities searched

- opeid: 8-digit OPE ID for institution
- unitid: Unit ID for institution
- schname: the search school name

``` {}
IDNameLinkRaw %>% 
  group_by(opeid) %>%
  summarise(count_unique = n_distinct(opeid)) %>% 
  arrange(., desc(count_unique))
```
Primary Key: both opeid & unitid are primary keys

```{}
summary(MostRecentRaw)

```
Data set holds 122 unique variables which are directly associated with out primary key combination
- opeid: 8-digit OPE ID for institution
- unitid: Unit ID for institution

Holds a variety of information about the school such as:
- school type
- location
- size
- admission rate
- standard entry test score distributions
- degree ypes awarded
- race 
- price
- etc..

Primary Key: both opeid & unitid are primary keys

``` {}
summary(TrendsUpToFinishRaw)
head(TrendsUpToFinishRaw)

```
Data set holds panel data for each university and key word index rating over time.
index ranged from 1-100 (this is relative to the school) indexes are not comparable directly
Primary Key: schid & keynum & monthorweek
** NOTE: Schid is not a unique value across similar panel data sets


``` {}
summary(TrendsUpToInter_1_Raw)
head(TrendsUpToInter_1_Raw)

```
``` {}
summary(TrendsUpToUphoenixRaw)
head(TrendsUpToUphoenixRaw)

```
TrendsUpToInter_1_Raw - 6 are identical sets buts account for different schools.
These tables are also the same structure but represent different school sets

- TrendsUpToUmRaw
- TrendsUpToUphoenixRaw
- TrendsUpToUtRaw
- TrendsUpToUtmbRaw
- TrendsUpToYorktowneRaw
- unioned Data sets 1-6 would make full set of TrendsUpToInter Data

Data set holds panel data for each university and key word index rating over time.
index ranged from 1-100 (this is relative to the school) indexes are not comparable directly
Primary Key: schid & keynum & monthorweek

## List of interesting variables
Cost Variables:

```{r, echo = TRUE}
df <- DataDictRaw %>% 
  select(NAME.OF.DATA.ELEMENT, VARIABLE.NAME) %>% 
  filter(DataDictRaw$dev.category == 'cost') %>% 
  show()
  
```


